### 2023-12-17 MSFT.Rmd
sarima

---
title: "The Impact of News Sentiment on MSFT"
author: "Ansam Zedan, Daniel Wullschleger"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
if (!require("vars")) install.packages("vars", dependencies=TRUE)
if (!require("tseries")) install.packages("tseries", dependencies=TRUE)
if (!require("quantmod")) install.packages("quantmod", dependencies=TRUE)
if (!require("PerformanceAnalytics")) install.packages("PerformanceAnalytics", dependencies=TRUE)
if (!require("forecast")) install.packages("forecast", dependencies=TRUE)
if (!require("rugarch")) install.packages("rugarch", dependencies=TRUE)
if (!require("tinytex")) install.packages("tinytex", dependencies=TRUE)
if (!require("zoo")) install.packages("zoo", dependencies=TRUE)
if (!require("ggplot2")) install.packages("ggplot2", dependencies=TRUE)

library(PerformanceAnalytics)
library(forecast)
library(rugarch)
library(vars)
library(tseries)
library(quantmod)
library(tinytex)
library(zoo)
library(ggplot2)
library(utils)
library(dplyr)
library(xts)
library(timeSeries)
```

# Introduction 
In recent years Microsoft has pushed and achieved great progress in a diverse portfolio of technologies (Microsoft, 2022) including also partnerships with AI research centers such as OpenAI. Alongside this development, its stock price has since reached new all-time highs and persists to be a major player in tech investments.

Inspired by this evolution, this paper provides an analysis and forecast of the stock price of Microsoft based on news sentiment data from subreddits such as r/worldnews, r/investing and r/stocks using a vector autoregressive (VAR) model.

Vector Autoregressive (VAR) processes are popular in economics for their flexibility and simplicity as models for multivariate time series data, forming a vector between variables that affect each other (Suharsono et al. 2017; Aptech 2021). The relationship between the stock prices and the news events will be analysed using a Granger causality testing, similarly as Bhowmik et al. (2022) proceeded when comparing multiple markets.

Hereby MSFT price data was downloaded through the Yahoo API in R and Reddit posts were scraped using Jupyter Notebooks. The python libraries nltk, textblob and wordlcoud were of support in the creation of the sentiment scores.

# Analysis:
First, we import the data and convert it to a time series so that we may plot it.
After importing and converting our dataset into a workable format, we observed noticeable noise in the average_polarity data. To address this, we applied a LOESS smoothing technique. This method allows us to smooth out short-term fluctuations and highlight longer-term trends in sentiment data, crucial for correlating with stock price movements. The choice of the smoothing parameter (span) was iteratively refined to balance detail and smoothness

```{r echo=FALSE, include=TRUE}
MSFT <- read.csv("data/final_data.csv")
MSFT$Index <- as.Date(MSFT$Index)
MSFT$MSFT.Adjusted <- as.numeric(as.character(MSFT$MSFT.Adjusted))
MSFT$average_polarity <- as.numeric(as.character(MSFT$average_polarity))
# MSFT$Index <- as.Date(MSFT$Index)
MSFT_ts <- zoo(MSFT$MSFT.Adjusted, order.by = MSFT$Index)

# plot(MSFT_ts)

events <- data.frame(
  date = as.Date(c("2023-01-23", "2022-11-30", "2023-06-29", "2023-04-12", "2023-11-15")),
  event = c(
    "Microsoft investing 10 billion dollars into OpenAI",
    "ChatGPT being released and created an AI-hype",
    "Microsoft releasing Bing Co-Pilot and Microsoft Co-Pilot",
    "The release of DALL-E 3 disrupting the image gen AI market",
    "Sam Altman leaving OpenAI in November 2023"
  )
)

# smooth polarity since it's very noisy
ix.na = is.na(MSFT$average_polarity)
problematic_values <- MSFT$average_polarity[ix.na]
MSFT$average_polarity[problematic_values] <- NA
# Replace with 0
MSFT$average_polarity[problematic_values] <- 0
MSFT$Index_numeric <- as.numeric(MSFT$Index - as.Date("2000-01-01"))
MSFT$smoothed_polarity_loess <- with(MSFT, loess(average_polarity ~ Index_numeric, span = 0.2)$fitted)

ggplot(data = MSFT, aes(x = Index)) +
  geom_line(aes(y = MSFT.Adjusted, color = "MSFT.Adjusted"), size = 1) +
  ylab("MSFT Adjusted Price") +

  geom_line(aes(y = smoothed_polarity_loess * 2000, color = "Sentiment Average Score"), size = 1) +
  ylab("Sentiment Average Score") +

  scale_y_continuous(
    name = "MSFT Adjusted Price",
    sec.axis = sec_axis(~./2000, name = "Sentiment Average Score"),
    limits = c(0, 400),
    breaks = seq(0, 400, by = 50)
  ) +

  scale_color_manual(
    values = c("MSFT.Adjusted" = "#00AFBB", "Sentiment Average Score" = "red"),
    labels = c("MSFT Adjusted Price", "Sentiment Average Score")
  ) +

  ggtitle("MSFT Adjusted Price and Sentiment Average Score Over Time") +
  theme_minimal()
```
We see a clear trend, which is an indicator for non-stationarity, so we do an Augmented Dickey-Fuller (ADF) Test to see whether the data is in fact non-stationary.

```{r echo=FALSE, include=TRUE}
adf.test(MSFT_ts) 
```
## Achieving Stationarity
We see that based on the p-value of 0.7051 the ADF Test suggests that the price data is non-stationary. So we a apply a log-transform with subsequent differencing to calculate continuous returns on MSFT. These should behave stationary.

```{r echo=FALSE, include=TRUE}
MSFT_returns <- na.omit(diff(log(MSFT_ts)))

plot(MSFT_returns)
adf.test(MSFT_returns)
```
## Creating a VAR Model using News sentiment
Now, that we deal with stationary data we can begin with our analysis.

We divide the dataset into a Train/Test Split, whereas November and December 2023 will serve as a test set for later predicitions.

First we create a time series of news sentiment data. Combined with the price returns, we can create a Vector Autoregression model. This will be our basis to check whether News are a causal driver for MSFT price returns.

We then do a coefficient test, to get more information on the statistical significance of the coefficients of our VAR Model.

```{r echo=FALSE, include=TRUE}
news_ts <- zoo(MSFT$average_polarity, order.by = MSFT$Index)

MSFT_returns_train <- window(MSFT_returns, start ="2022-01-04", end="2023-09-30")
MSFT_returns_test <- window(MSFT_returns, start="2023-10-01")

news_train <- window(news_ts, start ="2022-01-04", end="2023-09-30")
news_test <- window(news_ts, start="2023-10-01")

MSFT_VAR_Data <- cbind(MSFT_returns_train, news_train)
colnames(MSFT_VAR_Data) <- c("MSFT_Returns", "News_Sentiment")

MSFT_VAR <- VAR(MSFT_VAR_Data, ic="AIC", lag.max = 24)
coeftest(MSFT_VAR)
```
We see that almost none of the coefficients have a statistical significance as their p-values are all above 0.05. An exception is the second lag of MSFT returns. It seems to have a slight negative effect on the current price of MSFT. However, there is only small evidence.

## Causlaity testing
As a next step, we do a Granger causality test to confirm that our news sentiments are in fact not a causal driver for the MSFT stock price.

```{r echo=FALSE, include=TRUE}
causality(MSFT_VAR, cause="News_Sentiment")["Granger"] 
```
The p-value of 0.34 confirms that the gathered News sentiments are not a causal driver for the MSFT stock price. This leads us to reject our hypothesis.

## Forecasting
From a practical point of view, our VAR model may still be of some use when applied for forecasting purposes. To see if the inclusion of the news sentiment in our model has an improving effect on the models accuracy when predicting the price movements of MSFT we do some forecasting for our testing period where we include the news sentiment data of the testing period.

At the same time, we setup an ARMA model that is purely based on MSFTs previous price movements. It will serve as our benchmark.

```{r echo=FALSE, include=TRUE}
forecast_data <- data.frame(
  MSFT_Returns = rep(NA, length(news_test)),
  News_Sentiment = news_test
)

MSFT_VAR_pred <- predict(MSFT_VAR, newdata = forecast_data, n.ahead=43)
plot(MSFT_VAR_pred)
MSFT_VAR_pred_values <- MSFT_VAR_pred$fcst$MSFT_Returns

rmse_VAR <- sqrt(mean((MSFT_returns_test - MSFT_VAR_pred_values)^2))
cat("RMSE of VAR\t\t\t", rmse_VAR)
```

```{r echo=FALSE, include=TRUE}
MSFT_ARMA <- auto.arima(MSFT_returns_train)
MSFT_ARMA
MSFT_ARMA_pred <- forecast(MSFT_ARMA, level=95, h = length(news_test))
plot(MSFT_ARMA_pred)
```

```{r echo=FALSE, include=TRUE}
rmse_ARMA <- sqrt(mean((MSFT_returns_test - MSFT_ARMA_pred$mean)^2))
cat("RMSE of AR(2) Benchmark \t", rmse_ARMA)

```
We see that auto.arima has found that the order of the auto-regressive (AR) as well as the Moving Average (MA) parts are of order 0. Therefore, we essentially deal with a mean model as our benchmark. During forecasting, we also see that it uses a mean of 0 with a confidence band for all future values. This makes sense, as we deal with stationary data that commonly has a mean of 0.

While the VAR model reaches an RSME of 0.036 when compared to the truth values of our test set, the mean model benchmark reaches an RSME of 0.014. This means, our more sophisticated VAR model actually performs worse than the more simple benchmark that is purely based on the mean.

# Seasonal Adjustment

### STL Decomposition

After many trials of extracting seasonality by using the stl and decompose method and because the stock market data is not consistent because of holidays and the market is not open on a daily basis, here is another method to approach this problem using Fourier Transform:

```{r echo=FALSE, include=TRUE}
# Number of observations in the ime series
n <- length(MSFT_returns)

fourier <- data.frame(frequency = 1:(n/2), 
                      spectrum = abs(fft(coredata(MSFT_returns)))[1:(n/2)])

# Plot the Fourier Transform spectrum - can remove later
ggplot(fourier, aes(x = frequency, y = spectrum)) +
  geom_line() +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(title = "Fourier Transform of MSFT Adjusted Prices", 
       x = "Frequency", 
       y = "Amplitude")
```

We norice in this plot the presence of multiple peaks suggesting that there are several cyclical patterns in the data. This can happen if there are multiple underlying seasonalities or other forms of cyclical behavior. But there is not a single, dominant low-frequency component that clearly stands out. This suggests that the data might not have a strong trend or annual seasonality, or it may be masked by the noise. The problem is if we attempt to model all these frequencies, there's a risk of overfitting your time series model to the noise rather than to actual, meaningful seasonal patterns.

### Reevaluating seasonality and trend
```{r echo=FALSE, include=TRUE}
# Convert to ts object, starting from the first date, with a frequency of 365 (daily data)
start_date <- as.numeric(format(index(MSFT_returns)[1], "%Y")) + (as.numeric(format(index(MSFT_returns)[1], "%j")) - 1) / 365
MSFT_returns_ts <- ts(MSFT_returns, start = start_date, frequency = 365)

# Conduct a spectral analysis to identify significant frequencies
spectrum_analysis <- spectrum(MSFT_returns_ts, spans = c(3, 5), plot = FALSE)

# Identify peaks in the spectral density
peak_frequencies <- which(diff(sign(diff(spectrum_analysis$spec))) == -2) + 1
significant_frequencies <- spectrum_analysis$freq[peak_frequencies]
significant_spectrum <- spectrum_analysis$spec[peak_frequencies]

plot(spectrum_analysis$freq, spectrum_analysis$spec, type = 'l', xlab = 'Frequency', ylab = 'Spectral Density')
points(significant_frequencies, significant_spectrum, col = 'red', pch = 19)
```

```{r echo=FALSE, include=TRUE}
n <- length(MSFT_returns)
# Extract seasonal components using Fourier series
future_fourier_terms <- fourier(MSFT_returns_ts, K = length(significant_frequencies), h = 60) # 60 forecast horizon - 2 months

# Model the seasonality by including Fourier terms as external regressors
MSFT_seasonal_model <- auto.arima(MSFT_returns_ts, xreg = fourier_terms, seasonal = FALSE)
summary(MSFT_seasonal_model)
```
The best-fit model for the series is an ARIMA(2,0,1). This model includes autoregressive (AR) terms, moving average (MA) terms, and various regression coefficients for exogenous variables. The model's log-likelihood is 1259.22, and it exhibits a reasonable AIC value, which suggest a good balance between model fit and complexity. The BIC value is also indicative of a well-fitting model. The training set error measures indicate that the model performs well on the training data, with a small mean error (ME) and low root mean squared error (RMSE) and mean absolute error (MAE). However, the MPE and MAPE are relatively high, suggesting some room for improvement in capturing certain patterns in the data. The MASE is 0.6607, and the autocorrelation of residuals (ACF1) is -0.0065.

```{r echo=FALSE, include=TRUE}
# Fit the time series model and make forecasts
MSFT_forecast <- forecast(MSFT_seasonal_model, xreg = future_fourier_terms, h = 60)
plot(MSFT_forecast)
lines(MSFT_returns_test, col = 'red')
```
Calculate RMSE:
```{r echo=FALSE, include=TRUE}
actuals <- window(MSFT_returns_ts, start = c(2023, 118))
rmse <- sqrt(mean((actuals - MSFT_forecast$mean)^2, na.rm = TRUE))
cat("RMSE for Model with Fourier Terms: ", rmse)
```

# Conclusion
To be added
```{r echo=FALSE, include=TRUE}

```

# Bibliography:

Agus Shuarsono, Auliya Aziza and Wara Pramesti (2017); Comparison of vector autoregressive (VAR) and vector error correction models (VECM) for index of ASEAN stock price. DOI 10.1063/1.5016666

Aptech, Introduction to the Fundamentals of Vector Autoregressive Models (2021). https://www.aptech.com/blog/introduction-to-the-fundamentals-of-vector-autoregressive-models/

Microsoft, 2022 a look back at a year of accelerating progress in AI.
https://www.microsoft.com/en-us/research/blog/2022-a-look-back-at-a-year-of-accelerating-progress-in-ai/

Roni Bhowmik, Gouranga Chandra Debnath, Nitai Chandra Debnath, Shouyang Wang (2022); Emerging stock market reactions to shocks during various crisis periods 2022). DOI 10.1371/journal.pone.0272450